---
title: "Assignment_2"
author: "Vivekanand Kulkarni 19200231"
date: "08/11/2019"
output:
  html_document: default
  pdf_document: default
---
## 1 & 2)
Load in the data as an object called DublinAirport. Notice that you have to skip the
first 19 lines contained in the file before beginning to read data.

```{r}
DublinAirport=read.csv("V:\\Study\\R-Programming\\Assignments\\Assignment_2\\mly532.csv", header = TRUE,sep = ",",skip = 19)
str(DublinAirport)
```

## 3) Transform the Column month
Used factor to assign the month columns with month names. And also made use of the built-in constant month.name which is a vector of 12 months.
```{r}
DublinAirport$month<-factor(DublinAirport$month, levels = c('1', "2","3","4","5","6","7","8","9","10","11","12"),labels = month.name)
head(DublinAirport)
```

## 4)
Use the aggregate function to compute which month has on average the highest and
the lowest Precipitation Amount.
```{r}
Mean_rain=aggregate(DublinAirport$rain,list(DublinAirport$month),mean)
names(Mean_rain)[1]<-"Months"
names(Mean_rain)[2]<-"Mean of each month"
Mean_rain_sorted<- order(Mean_rain[,2],decreasing = TRUE)
print(paste0("The month with Highest rain Precipitation Amount is:",month.name[Mean_rain_sorted[1]]))
print(paste0("The month with Lowest rain Precipitation Amount is:",month.name[Mean_rain_sorted[12]]))
```
## 5)
Creating a new column which contains a factor indicating the season:
Winter: December, January, February,
Spring: March, April, May,
Summer: June, July, August,
Autumn: September, October, November
```{r}
DublinAirport$Season<-factor(DublinAirport$month ,levels =c("January", "February","March","April","May","June","July","August","September","October","November","December"),labels=c("Winter","Winter", "Spring","Spring","Spring", "Summer","Summer","Summer","Autumn","Autumn","Autumn","Winter"))

head(DublinAirport)
```
## 6)
Assign to the DublinAiport object the classes WeatherData and data.frame.
```{r}
class(DublinAirport)<-c('WeatherData','data.frame')
class(DublinAirport)
```
## 7)
Write an S3 summary method for an object of class WeatherData which produces
the following statistical summaries the rain, maxtp, mintp, maxgt variables split by
season: mean, standard deviation, minimum, maximum. Ignore the missing values in
the calculations.
Test your function on the DublinAirport data set and comment your fndings.
```{r}
summary.WeatherData<-function(x)
{
mean_rain <- aggregate(rain~Season, data = x, mean,na.rm= TRUE)
names(mean_rain)[1]<-"Mean per season"
max_rain <- aggregate(rain~Season, data = x,max,na.rm= TRUE)
names(max_rain)[1]<-"Max  per season"
sd_rain<- aggregate(rain~Season, data=x,sd,na.rm= TRUE)
names(sd_rain)[1]<-"Standard Deviation  per season"
min_rain<- aggregate(rain~Season, data=x,min,na.rm=TRUE)
names(min_rain)[1]<-"Min  per season"

print(mean_rain)
print(max_rain)
print(sd_rain)
print(min_rain)

mean_maxtp <- aggregate(maxtp~Season, data = x, mean,na.rm= TRUE)
names(mean_maxtp)[1]<-"Mean  per season"
max_maxtp <- aggregate(maxtp~Season, data = x,max,na.rm= TRUE)
names(max_maxtp)[1]<-"Max  per season"
sd_maxtp <- aggregate(maxtp~Season, data=x,sd,na.rm= TRUE)
names(sd_maxtp)[1]<-"Standard Deviation  per season"
min_maxtp <- aggregate(maxtp~Season, data=x,min,na.rm=TRUE)
names(min_maxtp)[1]<-"Min  per season"


print(mean_maxtp)
print(max_maxtp)
print(sd_maxtp)
print(min_maxtp)

mean_mintp <- aggregate(mintp~Season, data = x, mean,na.rm= TRUE)
names(mean_mintp)[1]<-"Mean  per season"
max_mintp <- aggregate(mintp~Season, data = x,max,na.rm= TRUE)
names(max_mintp)[1]<-"Max  per season"
sd_mintp <- aggregate(mintp~Season, data=x,sd,na.rm= TRUE)
names(sd_mintp)[1]<-"Standard Deviation per season"
min_mintp <- aggregate(mintp~Season, data=x,min,na.rm=TRUE)
names(min_mintp)[1]<-"Min  per season"


print(mean_mintp)
print(max_mintp)
print(sd_mintp)
print(min_mintp)

mean_maxgt <- aggregate(maxgt~Season, data = x, mean,na.rm= TRUE)
names(mean_maxgt)[1]<-"Mean  per season"
max_maxgt <- aggregate(maxgt~Season, data = x,max,na.rm= TRUE)
names(max_maxgt)[1]<-"Max  per season"
sd_maxgt<- aggregate(maxgt~Season, data=x,sd,na.rm= TRUE)
names(sd_maxgt)[1]<-"Standard Deviation per season"
min_maxgt<- aggregate(maxgt~Season, data=x,min,na.rm=TRUE)
names(min_maxgt)[1]<-"Min  per season"

print(mean_maxgt)
print(max_maxgt)
print(sd_maxgt)
print(min_maxgt)
}
summary(DublinAirport)
```
Created an S3 plot method for the class WeatherData that produces the following plots.
(1) Plot of the monthly Air Temperature (C) (maxtp, mintp).
(2) Plot of the Precipitation Amount (mm) (rain).
(3) Plot of the Highest Gust (knot) (maxgt).
The user must be able to decide which years to plot. By default it will use the
data from 2015 until 2018.
The user must be able to decide which plot to draw (i.e, only one of the three,
two of the three, or all three plots). By default the function will create all three
plots.
The plots must be on a single panel.
The plots must have meaningful labels and/or titles, and a legend if needed.
Test your function on the DublinAirport data set.
```{r}
DublinAirport<- na.omit(DublinAirport)
library(ggplot2)
library(gridExtra)
library(dplyr)
plot.WeatherData <- function(Dframe, yearfrom = 2015, yearto = 2018, number_of_plots = 3){
  
  # Selecting the data based on the range of years inputted to the function
  plot_data <- Dframe[Dframe$year >= yearfrom & Dframe$year <= yearto, ]

  # Plot of the monthly Air Temperature (C) (maxtp, mintp).  
  plot1 <- ggplot(plot_data, aes(x = year, y = maxtp, colour = month)) +
  geom_line(position = "jitter") + 
  ggtitle(" Plot of Air Temperature (C) (maxtp, mintp).") + 
  xlab("Years") + 
  ylab("Maximum and Minimum Temperatures") +
  theme(axis.text.x = element_text(angle = 90, face = "bold", size = 8)) +
  geom_line(aes(x = year, y = mintp, colour = month),position = "jitter") 
  
  # Plot of the Precipitation Amount (mm) (rain). 
  plot2 <- ggplot(plot_data, aes(x = year, y = rain, colour = month)) + 
  geom_line(position = "jitter") +
  ggtitle(" Plot of Precipitation Amount (mm) (rain).") + 
  xlab("Years") + 
  ylab("Precipitation Amount (mm) in terms of Rain") + 
  theme(axis.text.x = element_text(angle = 90, face = "bold", size = 8))

  # Plot of the Highest Gust (knot) (maxgt). 
  plot3 <- ggplot(plot_data, aes(x = year, y = maxgt, colour = month  )) + 
  geom_line(position = "jitter") +
  ggtitle("Highest Gust Plot(knot) (maxgt).") + 
  xlab("Year") + 
  ylab("Highest Gust (knot)") + 
  theme(axis.text.x = element_text(angle = 90, face = "bold", size = 8))
  
  # Comparing the number_of_plots variable, and retrieving according to the request.
  if(number_of_plots == 1)
  {
    #Display plot1 if number_of_plots is 1
    plot1
  }
  else if(number_of_plots == 2)
  {
    #Display plot1 & plot2 if number_of_plots is 2
    require(gridExtra)
    grid.arrange(plot1, plot2,nrow =1, ncol = 2)
  }
  else if(number_of_plots == 3)
  {
    # By Default all the 3 plots are to be Displayed.
    # Display plot1, plot2 & plot3 if the number_of_plots is 3
    require(gridExtra)
    grid.arrange(plot1, plot2, plot3, nrow =2, ncol = 2)
  }
}

# The plot function is called and defined for the WeatherData Class
# Arguments to plot() functions are 
# Dframe --> The Data Frame, in this case DublinAirport
# yearfrom --> The year to begin with
# yearto --> Till which year the data should be selected.
# number_of_plots --> Number of plots that are to be  displayed among the 3 available plots. By default all 3 plots are displayed.

# For the demonstration purpose, Dframe = DublinAirport,  yearfrom = 2015, yearto = 2018, number_of_plots = 3
plot(DublinAirport, 2015,2018,3)
```

